---
title: "Open University Learning Analytics Dataset"
author: "Fanita Rodrigues"
date: "4/24/2018"
output: 
 md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

 Open University Learning Analytics Dataset (OULAD) contains data about 
    courses, students and their interactions with Virtual Learning 
    Environment (VLE) for seven selected courses (called modules). Presentations
    of courses start in February and October - they are marked by "B" and "J" 
    respectively. The dataset consists of tables connected using unique 
    identifiers. Dataset is stored in several csv files. We start off the analysis by declaring various packages that we will use.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(data.table)
library(lubridate)
library(plotly)
```

## Exploratory Data Analysis

We can now start reading in the csv files and explore the data within. The `tidyverse` package is ideally suited for all data wrangling operations. Since one of the csv files (`studentVle.csv`) is 432 MB in size, we will utilize the `fread()` function from the `data.table` package to speed up data ingestion.

```{r, warning=FALSE, message=FALSE}
assess <- read_csv("assessments.csv")
courses <- read_csv("courses.csv")
stuAssess <- read_csv("studentAssessment.csv")
stuInfo <- read_csv("studentInfo.csv")
stuReg <- read_csv("studentRegistration.csv")
stuVle <- fread("studentVle.csv")
vle <- read_csv("vle.csv")
```

## Biggest Problem Area
Let's start off by exploring our students.

```{r}
glimpse(stuInfo)
colSums(is.na(stuInfo))
```

As an educator, I would be most concerned about students who fail in the courses or those who withdraw.
```{r}
table(stuInfo$final_result)
```

As we can see `r round(10156/nrow(stuInfo),2)*100`% of the students withdraw from the courses. Further, `r round(7052/nrow(stuInfo),2)*100`% of the students fail in courses. We can now dive a bit deeper into these students who fail or withdraw from courses.

```{r}
stuInfo$gender <- as.factor(stuInfo$gender)
stuInfo$final_result <- as.factor(stuInfo$final_result)
stuWF <- stuInfo %>% filter(final_result=='Withdrawn'|final_result=='Fail')

ggplot(stuWF, aes(x=final_result, fill=gender))+geom_bar()
```

Are there any particular modules that these students are facing problems in?
```{r}
stuInfo$code_module <- as.factor(stuInfo$code_module)
ggplot(stuWF, aes(x=code_module, fill=final_result))+geom_bar()
```

What about the level of education?

```{r}
stuInfo$highest_education <- as.factor(stuInfo$highest_education)
ggplot(stuWF, aes(x=highest_education, fill=final_result))+geom_bar()
```

After how many days are students withdrawing? In order to answer this, let's combine the dates of registration and unregistration from the `stuReg` table using the `left_join()` function from `dplyr`.

```{r,warning=FALSE,error=FALSE}
stuWF <- stuWF %>% left_join(stuReg) %>% dplyr::select(names(stuWF), date_registration, date_unregistration)

stuWF$date_unregistration[stuWF$date_unregistration=="?"] <- NA
stuWF$date_unregistration <- as.numeric(stuWF$date_unregistration)
p <- ggplot(stuWF, aes(x=date_unregistration, fill = cut(date_unregistration, 100)))+geom_histogram(show.legend = FALSE,binwidth = 20)
p
```
 
How many students leave even before the course starts?
```{r}
stuWF %>% filter(date_unregistration<0) %>% count()
```

Let's see the students who leave the course after it commences.

```{r}
stuW <- stuWF %>% filter(date_unregistration>0)
ggplot(stuW, aes(x=date_unregistration, fill = cut(date_unregistration, 100)))+geom_histogram(show.legend = FALSE,binwidth = 20)
```

A lot of students appear to be leaving within the first 2 months. The number of students leaving during this duration are `r stuW %>% filter(date_unregistration<=60) %>% count()`.